The Strong Gravitational Lens Finding Challenge

### 2017-05-03 paper.tex
Edit

@@

The challenge was in fact two independent challenges that could be entered independently.  One was designed to mimic a single band of a future imaging data set from a satellite survey such as Euclid\footnote{https://www.euclid-ec.org/} \citep{***}.  The other was designed to mimic ground based data with multiple bands, roughly modeled on the Kilo-Degree Survey (KiDS)\footnote{http://kids.strw.leidenuniv.nl/} \citep{***} .    In neither case were the simulated images meant to precisely mock these surveys, but only used them as general guide as to set noise levels, pixel sizes, sensitivities and other parameters.

@@

For lensed sources we use sources from the Hubble Ultra Deep Field (UDF) that have been decomposed into shapelet functions to remove noise.  This is the same set of images as used 
in ***.   The source is placed within **** of the center of the lenses caustic as found in the initial ray tracing.  This is a somewhat arbitrary length designed to be a compromise between producing only very clear strong lenses because all the sources are right in the center of the caustic and making the proccess  inefficiently because most of the sources are too far away from the caustic to produce clear lenses.  If the source positions where taken completely at random the fraction of clear lenses would be very low.

@@

For the training sets all these products were provided, but for the test sets only the final images were provided to participants although all the information was stored for analysing the entries.

@@

In the test image set some actual KiDS images were used also.  Here a random mass distribution and source was used to produce a lensed source that was added directly to the KiDS image after aplying the PSF.  No additional noise was added.

@@

By chance one of the original KiDS images appears to have been a lens.  When an additional lensed source was added this made a double lens or "jockpot" lens.

@@

\begin{figure}
 \includegraphics[width=\columnwidth]{figures/arcmethod.pdf}
 \caption{ (GAHEC IRAP) From top-left to botton right, 1) a simulated arc extracted from SL challenge in which an tunned Arcfinder selects 3 candidates (green circles), 2) the smoothed image on which pixelwise elongation is computed, 3) the resulting elongated pixels after thresholding, 4) the set of pixels selected for the computation of arc candidate properties. }
 \label{fig:Cabanac}
\end{figure}

Arcfinder \citep{2006astro.ph..6757A,2007A&A...461..813C,2012ApJ...749...38M} is a fast linear method that computes a pixelwise elongation parameter (ratio of first-order moments in a n-pix window oriented in proper reference frame) for all pixels of mexican-hat-smoothed FITS images. Arcfinder then extract contiguous pixels above a given background and computes the candidate arcs length, width, area, radius of curvature and peak surface brightness. A final thresholding is set to maximize purity over completeness on a few typical arcs of the dataset.
For the current SL challenge, arcfinder was tunned to detect long and narrow arcs, and was optimized on a subset of 1000 simulated images with a grid covering a range of elongation windows and arc areas.  A python wrapper allows users to change parameters in a flexible way and run the arcfinder C code as a linux line command. Arcfinder took a couple of hours to run on the entire dataset with some overheads due to the dataset format. The code is publicly available at https://github.com/rcabanac/arcfinder.

@@

The lens detector is based on a convolutional neural network (CNN), trained with the provided training datasets. The CNN is implemented in Python, using the TensorFlow framework\footnote{http://www.tensorflow.org/}. Both ground multichannel and space monochannel image classifiers have the exact same CNN architecture.

@@

A Support Vector Machine (SVM) is a supervised machine learning method which uses labelled training data to determine a classification model (see e.g., \citet{vapnik79estimation}, \citet{Cortes1995} and \citet{Burges1998}). A preprocessing stage first extracts a set of useful features from input samples, before projecting each sample as a vector in multidimensional space. The model then separates classes of data by maximising the margin between a defining hyperplane and a set of so-called support-vectors at the inner edge of each class. The process of classification is computationally inexpensive since optimisation depends only on the dot products of the support vector subset. Feature extraction, however,  requires both an extensive exploration of the feature space during the development of a model, and potentially intensive computer resources in order to transform the original samples. Our full method is described in detail in Hartley et. al., in prep., and was developed using the Python scikit-learn and scikit-image packages \citep{scikit-learn,scikit-image}.

@@

Figures~\ref{fig:einstein_space} and \ref{fig:einstein_ground} shows the AUROC, TPR$_0$, TPR$_{10}$ and fraction of lenses as a function of a lower cutoff on the Einstein radius (area).  There is also a vertical dotted line that indicates where no more than 100 lenses in the test sample had Einstein radii larger than this mark.  Beyond this point one should be suspicius of small number statistics.  When deriving the distribution of Einstein radii from data these curves would need to be used to correct for detection bias.  It can be seen the CMU-DeepLens-ResNet, Manchester1, Manchester SVM and GAMOCLASS obtain significantly higher TPR$_0$ and  TPR$_{10}$ for larger Einstein radii.  Manchester1 is the human inspection method.  In some cases these are above 50\% of the lenses that meet the criterion.  Remember that many of the so called lenses are very dim or there is no significant arc because of the source position is well outside the caustic.  If an additional require was placed on the definition of a lens, such as the brightness of the arc being above a threshold, the TPRs would go up.

Figures~\ref{fig:flux_space} and \ref{fig:flux_ground} are the same except that the flux in the lensed image is used as the threshold.  We count only the flux in pixels with flux over one $\sigma$ of the background.  In some cases one can see an abrupt rise in the TPRs at some flux threshold.  CMU-DeepLens-ResNet in particular reaches a TPR$_0$ above 75\% for the bightest $\sim$ 10\% of the lenses.

@@

Figures~\ref{fig:flux_contrast_space} and \ref{fig:flux_contrast_ground} investigate how the flux contrast between the foreground objects and the lensed source affects the classifiers.  Interestingly some method's TPRs go up with this quantity and some go down.  \red{explaination?}

>>>>

Arcfinder\citep{2007A&A...472..341S}

\citet{2006astro.ph..6757A} arc finder used in the CFHT \citep{2007A&A...461..813C,2012ApJ...749...38M}

Some finders compare each image with a lens model \citep{2009ApJ...694..924M,2010A&A...517A..25S}.

\citep{2014A&A...566A..63J,2016A&A...592A..75P}

@@

The challenge was in fact two independent challenges that could be entered independently.  One was designed to mimic a single band of a future imaging data set from a satellite survey such as Euclid\footnote{https://www.euclid-ec.org/} \citep{***}.  The other was designed to mimic ground based data with multiple bands, roughly modeled on the Kilo-Degree Survey (KiDS)\footnote{http://kids.strw.leidenuniv.nl/} \citep{2013ExA....35...25D} .    In neither case were the simulated images meant to precisely mock these surveys, but only used them as general guide as to set noise levels, pixel sizes, sensitivities and other parameters.

@@

For lensed sources we use sources from the Hubble Ultra Deep Field (UDF) that have been decomposed into shapelet functions to remove noise.  This is the same set of images as used 
in ****.   The source is placed within **** of the center of the lenses caustic as found in the initial ray tracing.  This is a somewhat arbitrary length designed to be a compromise between producing only very clear strong lenses because all the sources are right in the center of the caustic and making the process  inefficiently because most of the sources are too far away from the caustic to produce clear lenses.  If the source positions where taken completely at random the fraction of clear lenses would be very low.

@@

For the training sets all these products were provided, but for the test sets only the final images were provided to participants although all the information was stored for analyzing the entries.

@@

In the test image set some actual KiDS images were used also.  Here a random mass distribution and source was used to produce a lensed source that was added directly to the KiDS image after applying the PSF.  No additional noise was added.

@@

By chance one of the original KiDS images appears to have been a lens.  When an additional lensed source was added this made a double lens or "jackpot" lens.

@@

\begin{figure}
 \includegraphics[width=\columnwidth]{figures/arcmethod.pdf}
 \caption{ (GAHEC IRAP) From top-left to bottom right, 1) a simulated arc extracted from SL challenge in which an tuned Arcfinder selects 3 candidates (green circles), 2) the smoothed image on which pixel wise elongation is computed, 3) the resulting elongated pixels after threshold, 4) the set of pixels selected for the computation of arc candidate properties. }
 \label{fig:Cabanac}
\end{figure}

Arcfinder \citep{astronaut.ph..6757A,2007A&A...461..813C,2012ApJ...749...38M} is a fast linear method that computes a pixel wise elongation parameter (ratio of first-order moments in a n-pix window oriented in proper reference frame) for all pixels of mexican-hat-smoothed FITS images. Arcfinder then extract contiguous pixels above a given background and computes the candidate arcs length, width, area, radius of curvature and peak surface brightness. A final thresholding is set to maximize purity over completeness on a few typical arcs of the data set.
For the current SL challenge, arcfinder was tuned to detect long and narrow arcs, and was optimized on a subset of 1000 simulated images with a grid covering a range of elongation windows and arc areas.  A python wrapper allows users to change parameters in a flexible way and run the arcfinder C code as a linux line command. Arcfinder took a couple of hours to run on the entire data set with some overheads due to the data set format. The code is publicly available at https://github.com/rcabanac/arcfinder.

@@

The lens detector is based on a convolutional neural network (CNN), trained with the provided training data sets. The CNN is implemented in Python, using the TensorFlow framework\footnote{http://www.tensorflow.org/}. Both ground multichannel and space monochannel image classifiers have the exact same CNN architecture.

@@

A Support Vector Machine (SVM) is a supervised machine learning method which uses labeled training data to determine a classification model (see e.g., \citet{vapnik79estimation}, \citet{Cortes1995} and \citet{Burges1998}). A preprocessing stage first extracts a set of useful features from input samples, before projecting each sample as a vector in multidimensional space. The model then separates classes of data by maximising the margin between a defining hyperplane and a set of so-called support-vectors at the inner edge of each class. The process of classification is computationally inexpensive since optimisation depends only on the dot products of the support vector subset. Feature extraction, however,  requires both an extensive exploration of the feature space during the development of a model, and potentially intensive computer resources in order to transform the original samples. Our full method is described in detail in Hartley et. al., in prep., and was developed using the Python scikit-learn and scikit-image packages \citep{scikit-learn,scikit-image}.

@@

Figures~\ref{fig:einstein_space} and \ref{fig:einstein_ground} shows the AUROC, TPR$_0$, TPR$_{10}$ and fraction of lenses as a function of a lower cutoff on the Einstein radius (area).  There is also a vertical dotted line that indicates where no more than 100 lenses in the test sample had Einstein radii larger than this mark.  Beyond this point one should be suspicious of small number statistics.  When deriving the distribution of Einstein radii from data these curves would need to be used to correct for detection bias.  It can be seen the CMU-DeepLens-ResNet, Manchester1, Manchester SVM and GAMOCLASS obtain significantly higher TPR$_0$ and  TPR$_{10}$ for larger Einstein radii.  Manchester1 is the human inspection method.  In some cases these are above 50\% of the lenses that meet the criterion.  Remember that many of the so called lenses are very dim or there is no significant arc because of the source position is well outside the caustic.  If an additional require was placed on the definition of a lens, such as the brightness of the arc being above a threshold, the TPRs would go up.

Figures~\ref{fig:flux_space} and \ref{fig:flux_ground} are the same except that the flux in the lensed image is used as the threshold.  We count only the flux in pixels with flux over one $\sigma$ of the background.  In some cases one can see an abrupt rise in the TPRs at some flux threshold.  CMU-DeepLens-ResNet in particular reaches a TPR$_0$ above 75\% for the brightest $\sim$ 10\% of the lenses.

@@

Figures~\ref{fig:flux_contrast_space} and \ref{fig:flux_contrast_ground} investigate how the flux contrast between the foreground objects and the lensed source affects the classifiers.  Interestingly some method's TPRs go up with this quantity and some go down.  \red{explanation?}
